Supplementary Figures

### 2019-05-18 supplementary_figures/supp2_filtering.Rmd
updated supp figures

```{r}
suppressPackageStartupMessages({
  library(ggplot2)
  library(cowplot)
  library(SingleCellExperiment)
  library(reshape2)
  library(ComplexHeatmap)
  library(knitr)
})
suppressMessages(devtools::load_all("../../pipComp/"))
source("../misc_functions.R")
data(datasets)

# we load the datasets
lf <- list.files("../datasets/", full.names=TRUE)
names(lf) <- gsub(".SCE.rds", "", basename(lf), fixed=TRUE)
SCEs <- lapply(lf, readRDS)
```

@@

```{r filt_more_ari, fig.width=8, fig.height=8.5}
res <- readRDS("../data/filtering_old.rds")
res$pcCells <- round(100*res$nbCells/datasets[as.character(res$dataset),"nbCells"],2)
# we restrict to analyses that gave a number of clusters close to reality:
res2 <- res[which(abs(res$nbClusters-datasets[as.character(res$dataset), "subpopulations"])<=1),]
p1 <- ggplot(res2, aes(x=pcCells, y=ARI, colour=norm)) + geom_point() + facet_wrap(~dataset, scale="free")

fres <- readRDS("../../resNew/filtering_intermediateSummary.rds")
pc <- fres$filtering$pcCells
pc <- pc[grep("doublet",row.names(pc),invert=T),]
row.names(pc) <- reduceMethodNames(row.names(pc),2)
ha <- HeatmapAnnotation(dataset=reduceMethodNames(colnames(pc),1))
shortDSNames <- c("mixology10x3cl"="mix3cl", "mixology10x5cl"="mix5cl", "Zhengmix4eq"="Zh4eq", "Zhengmix4uneq"="Zh4uneq", "Zhengmix8eq"="Zh8eq")
colnames(pc) <- paste(shortDSNames[reduceMethodNames(colnames(pc),1)], reduceMethodNames(colnames(pc),2))
p2 <- grid.grabExpr(draw(Heatmap(pc, name="% out", bottom_annotation=ha)))

plot_grid(p1, p2, labels=LETTERS[1:2], nrow=2)
```

@@

```{r}
sw <- fres$dimreduction$clust.avg.silwidth[[1]]
sw <- sw[,grep("doublet",colnames(sw))]
w <- grep("seuratvst",colnames(sw))
colnames(sw) <- reduceMethodNames(colnames(sw),2)

chm(list("standard Seurat norm"=sw[,-w], "sctransform"=sw[,w]), scale="row", scaleTitle="z-score", row_ann=data.frame(meanSilWidth=rowMeans(sw), dataset=reduceMethodNames(row.names(sw),1)), show_row_names=FALSE)
```

```{r fig.height=3}
r <- fres$dimreduction$PC.R2[which(fres$dimreduction$PC.R2$dim<=10),]
r <- r[grep("doublet",r$method),]
r <- r[grep("clust",r$method, invert=T),]
r <- aggregate(r$value, by=r[,c(1,3)], FUN=mean)
r <- t(cast2(r, dataset~method, value.var="x", fun.aggregate=mean))
w <- grep("seuratvst", row.names(r))
row.names(r) <- reduceMethodNames(row.names(r), 2)
chm(list("standard Seurat norm"=t(r[-w,]), "sctransform"=t(r[w,])), sameScale=TRUE, 
    scale="row", scaleTitle="z-score", value_format="%.2f")
```

@@

```{r}
res <- readRDS("../data/selection_old.rds")
res <- res[grep("2000$",res$sel),]
res <- res[grep("deviance|seurat.variance.standardized",res$sel),]
res$filt <- as.factor(res$filt)
levels(res$filt) <- c("coding only", "no mitochondrial", "no ribosomal", "all")
res2 <- res[which(abs(res$nbClusters-datasets[res$dataset,"subpopulations"])<=1),]
ll <- lapply(split(res2, res2$norm), FUN=function(x){
  m2 <- cast2(x, formula=filt~dataset, fun.aggregate=median, value.var="ARI")
})
chm(ll, scale="column", scaleTitle="z-score", sameScale=TRUE, value_format="%.2f", row_sortFn=function(x){ median(x,na.rm=T) }, cluster_columns=FALSE)
```

>>>>

```{r}
suppressPackageStartupMessages({
  library(ggplot2)
  library(cowplot)
  library(SingleCellExperiment)
  library(reshape2)
  library(ComplexHeatmap)
  library(knitr)
  library(circlize)
})
suppressMessages(devtools::load_all("../../pipComp/"))
source("../misc_functions.R")
data(datasets)
```

@@

```{r filt_more_ari, fig.width=8, fig.height=6.5}
res <- readRDS("../data/filtering_old.rds")
res$pcCells <- round(100*res$nbCells/datasets[as.character(res$dataset),"nbCells"],2)
# we restrict to analyses that gave a number of clusters close to reality:
res2 <- res[which(abs(res$nbClusters-datasets[as.character(res$dataset), "subpopulations"])<=1),]
p1 <- ggplot(res2, aes(x=pcCells, y=ARI, colour=norm)) + geom_point() + facet_wrap(~dataset, scale="free")

fres <- readRDS("../../resNew/filtering_intermediateSummary.rds")
pc <- fres$filtering$pcCells
pc <- pc[grep("doublet",row.names(pc),invert=T),]
row.names(pc) <- reduceMethodNames(row.names(pc),2)
ha <- HeatmapAnnotation(dataset=reduceMethodNames(colnames(pc),1))
# shortDSNames <- c("mixology10x3cl"="mix3cl", "mixology10x5cl"="mix5cl", "Zhengmix4eq"="Zh4eq", "Zhengmix4uneq"="Zh4uneq", "Zhengmix8eq"="Zh8eq")
# colnames(pc) <- paste(shortDSNames[reduceMethodNames(colnames(pc),1)], reduceMethodNames(colnames(pc),2))
# p2 <- grid.grabExpr(draw(Heatmap(pc, name="% out", bottom_annotation=ha)))

pal <- colorRamp2(c(0,1,5,20,max(pc)),c("#0000FFFF", "#B38BFFFF", "#FFFFFFFF", "#FF9E81FF", "#FF0000FF"))
p2 <- grid.grabExpr(draw(Heatmap(pc, name="% out", bottom_annotation=ha, col=pal, cluster_columns=FALSE, show_column_names=FALSE)))

plot_grid(p1, p2, labels=LETTERS[1:2], nrow=2, rel_heights=c(3,2))
```

@@

```{r, fig.height=5}
sw <- fres$dimreduction$clust.avg.silwidth[[1]]
sw <- sw[,grep("doublet",colnames(sw))]
w <- grep("seuratvst",colnames(sw))
colnames(sw) <- reduceMethodNames(colnames(sw),2)
sw <- sw[order(reduceMethodNames(row.names(sw),1), rowMeans(sw)),]
ra <- data.frame(meanSilWidth=rowMeans(sw), dataset=reduceMethodNames(row.names(sw),1))
chm(list("standard Seurat norm"=sw[,-w], "sctransform"=sw[,w]), scale="row", scaleVar=FALSE, scaleTitle="meanSilWidth\ndifference", sameScale=TRUE, row_ann=ra, show_row_names=FALSE, cluster_rows=FALSE)
```

```{r fig.height=3.5, fig.width=6}
r <- fres$dimreduction$PC.R2[which(fres$dimreduction$PC.R2$dim<=10),]
r <- r[grep("doublet",r$method),]
r <- r[grep("clust",r$method, invert=T),]
r <- aggregate(r$value, by=r[,c(1,3)], FUN=mean)
r <- t(cast2(r, dataset~method, value.var="x", fun.aggregate=mean))
w <- grep("seuratvst", row.names(r))
row.names(r) <- reduceMethodNames(row.names(r), 2)
chm(list("standard Seurat norm"=t(r[-w,]), "sctransform"=t(r[w,])), sameScale=TRUE, 
    scale="row", scaleTitle="z-score", value_format="%.2f")
```

@@

```{r}
res <- readRDS("../data/selection_old.rds")
res <- res[grep("deviance|seurat.variance.standardized",res$sel),]
res$filt <- as.factor(res$filt)
levels(res$filt) <- c("coding only", "no mitochondrial", "no ribosomal", "all")
res2 <- res[which(abs(res$nbClusters-datasets[res$dataset,"subpopulations"])<=1),]
ll <- lapply(split(res2, res2$norm), FUN=function(x){
  m2 <- cast2(x, formula=filt~dataset, fun.aggregate=median, value.var="ARI")
})
chm(ll, scale="column", scaleTitle="z-score", sameScale=TRUE, value_format="%.2f", row_sortFn=function(x){ median(x,na.rm=T) }, cluster_columns=FALSE)
```

### 2019-05-18 supplementary_figures/supp3_normalization.Rmd
updated supp figures

```{r norm_r2, fig.width=8, fig.height=9}
res <- readRDS("../../resNew/norm_intermediateSummary.rds")
r2 <- res$dimreduction$PC.R2
r2$norm <- gsub("norm.","",reduceMethodNames(r2$method,3),fixed=T)
r2 <- r2[which(r2$dim <= 10),]
r2$R.on.Clusters <- sqrt(r2$value)
r3 <- cast2(r2, formula=norm~dataset, value.var="R.on.Clusters")
r2m <- dplyr::summarise(group_by(r2,dataset), meanR=mean(R.on.Clusters))
p1 <- ggplot(r2, aes(norm, R.on.Clusters, fill=norm)) + geom_violin() + facet_wrap(~dataset, nrow=1) + coord_flip() + theme(legend.position="none") + stat_summary(fun.y="mean", geom="point") + geom_hline(data=r2m, aes(yintercept=meanR)) + ggtitle("Proportion of the top dimensions explained by real clusters") + ylab("sqrt(R^2) of dimension ~ real cluster")

pc1.covar <- lapply( res$dimreduction$PC1.covar[2:3], FUN=function(x){
  x <- aggregate(x, by=list(reduceMethodNames(row.names(x), 3)), FUN=median)
  row.names(x) <- x[,1]
  x[,1] <- NULL
  x
})
names(pc1.covar) <- c("Correlation with log10-counts", "Correlation with nb features")
p2 <- grid.grabExpr(draw(chm(pc1.covar, sameScale=TRUE, scaleTitle="cor", value_format="%.2f", row_sortFn=median)))
plot_grid(p2, p1, labels=LETTERS[1:2], nrow=2)
```

### Supplementary Figure `r getFigNb()`
**A:** Proportion of the variance in the first 10 PCA components that is explained by real subpopulations. The square-root of R-squared values is plotted for ease of visualization.
**B:** Correlation of the first component with library size (left) or with the number of detected features (right).

@@

```{r norm_silwidth, fig.width=8, fig.height=7}
res <- readRDS("../../resNew/norm_intermediateSummary.rds")
sw <- res$dimreduction$clust.avg.silwidth[[1]]
hma <- HeatmapAnnotation(filtering=reduceMethodNames(colnames(sw),2), col=list(filtering=c(deffilter="#CC6677",lenientfilter="#4477AA",clustFilterStringent="#DDCC77")))
colnames(sw) <- gsub("norm.","",reduceMethodNames(colnames(sw), 3),fixed=T)
chm(list("Cluster mean silhouette width"=sw), scaleTitle="z-score", scale="row", row_ann=data.frame(meanSilWidth=rowMeans(sw), dataset=reduceMethodNames(row.names(sw),1)), show_row_names=FALSE, top_annotation=hma)
```

>>>>

```{r norm_r2, fig.width=8, fig.height=9}
res <- readRDS("../../resNew/norm2_intermediateSummary.rds")
r2 <- res$dimreduction$PC.R2
r2$norm <- gsub("norm.","",reduceMethodNames(r2$method,3),fixed=T)
r2 <- r2[which(r2$dim <= 10),]
r2$R.on.Clusters <- sqrt(r2$value)
r3 <- cast2(r2, formula=norm~dataset, value.var="R.on.Clusters")
r2m <- dplyr::summarise(group_by(r2,dataset), meanR=mean(R.on.Clusters))
p1 <- ggplot(r2, aes(norm, R.on.Clusters, fill=norm)) + geom_violin() + facet_wrap(~dataset, nrow=1) + coord_flip() + theme(legend.position="none") + stat_summary(fun.y="mean", geom="point") + geom_hline(data=r2m, aes(yintercept=meanR)) + ggtitle("Proportion of the top dimensions explained by real clusters") + ylab("sqrt(R^2) of dimension ~ real subpopulation")

pc1.covar <- lapply( res$dimreduction$PC1.covarR[2:3], FUN=function(x){
  x <- aggregate(x, by=list(reduceMethodNames(row.names(x), 3)), FUN=median)
  row.names(x) <- x[,1]
  x[,1] <- NULL
  x
})
names(pc1.covar) <- c("Correlation with log10-counts", "Correlation with nb features")
p2 <- grid.grabExpr(draw(chm(pc1.covar, sameScale=TRUE, scaleTitle="cor", value_format="%.2f", row_sortFn=median)))
plot_grid(p2, p1, labels=LETTERS[1:2], nrow=2)
```

### Supplementary Figure `r getFigNb()`
**A:** Correlation between library size (left) or the number of detected features (right) and the residuals of a linear regression of the first principal component on the subpopulation.
**B:** Proportion of the variance in the first 10 PCA components that is explained by real subpopulations. The square-root of R-squared values is plotted for ease of visualization.

@@

```{r norm_silwidth, fig.width=8, fig.height=7}
res <- readRDS("../../resNew/norm2_intermediateSummary.rds")
sw <- res$dimreduction$clust.avg.silwidth[[1]]
colnames(sw) <- gsub("norm.","",reduceMethodNames(colnames(sw), 3),fixed=T)
sw <- sw[order(reduceMethodNames(row.names(sw),1), rowMeans(sw)),]
ra <- data.frame(meanSilWidth=rowMeans(sw), dataset=reduceMethodNames(row.names(sw),1))
chm(list("Cluster mean silhouette width"=sw), scaleTitle="row z-score", scale="row", row_ann=ra, show_row_names=FALSE, col_sortFn=mean, cluster_rows=FALSE)
```

### 2019-05-18 supplementary_figures/supp_figures.tex
updated supp figures

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1.5cm]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Supplementary Figures},
            pdfauthor={Pierre-Luc Germain},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{natbib}
\bibliographystyle{plainnat}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

  \title{Supplementary Figures}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
    \author{Pierre-Luc Germain}
    \preauthor{\centering\large\emph}
  \postauthor{\par}
      \predate{\centering\large\emph}
  \postdate{\par}
    \date{12 Mai, 2019}

@@

\vfill

@@

\includegraphics{supp_figures_files/figure-latex/misclass-1.pdf}
\vspace{1cm}
\includegraphics{supp_figures_files/figure-latex/unnamed-chunk-4-1.pdf}

\vfill

@@

@@

\includegraphics{supp_figures_files/figure-latex/unnamed-chunk-5-1.pdf}
\vspace{1cm}
\includegraphics{supp_figures_files/figure-latex/unnamed-chunk-6-1.pdf}

\vfill

Relationship between various cellular properties and the frequency of
cluster mis-assignment for the Zheng equal (A) or unequal (B) mixtures
of four cell types. See Supplementary Figure 4 for more information. The
only clear pattern is that cells with a high number of reads or features
tend to have a higher misclassification rate.

@@

\includegraphics{supp_figures_files/figure-latex/unnamed-chunk-7-1.pdf}

\vfill

Relationship between various cellular properties and the frequency of
cluster mis-assignment for the Zheng mixture of 8 cell types. See
Supplementary Figure 4 for more information.

>>>>

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1.5cm]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Supplementary Figures},
            pdfauthor={Pierre-Luc Germain},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{natbib}
\bibliographystyle{plainnat}
\usepackage{longtable,booktabs}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

  \title{Supplementary Figures}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
    \author{Pierre-Luc Germain}
    \preauthor{\centering\large\emph}
  \postauthor{\par}
      \predate{\centering\large\emph}
  \postdate{\par}
    \date{18 Mai, 2019}

@@

@@

\includegraphics{supp_figures_files/figure-latex/cellprops_cor-1.pdf}

@@

\includegraphics{supp_figures_files/figure-latex/misclass-1.pdf} \vfill
\includegraphics{supp_figures_files/figure-latex/unnamed-chunk-7-1.pdf}

@@

\includegraphics{supp_figures_files/figure-latex/unnamed-chunk-8-1.pdf}
\vfill
\includegraphics{supp_figures_files/figure-latex/unnamed-chunk-9-1.pdf}

Relationship between various cellular properties and the frequency of
cluster mis-assignment for the Zheng equal (A) or unequal (B) mixtures
of four cell types. See Supplementary Figure 5 for more information. The
only clear pattern is that cells with a high number of reads or features
tend to have a higher misclassification rate.

@@

\hypertarget{supplementary-figure-7}{%
\section{Supplementary Figure 7}\label{supplementary-figure-7}}

\includegraphics{supp_figures_files/figure-latex/unnamed-chunk-10-1.pdf}

\hypertarget{supplementary-figure-7-1}{%
\subsubsection{Supplementary Figure 7}\label{supplementary-figure-7-1}}

Relationship between various cellular properties and the frequency of
cluster mis-assignment for the Zheng mixture of 8 cell types. See
Supplementary Figure 5 for more information.

\hypertarget{supplementary-figure-8}{%
\section{Supplementary Figure 8}\label{supplementary-figure-8}}

\includegraphics{supp_figures_files/figure-latex/filt_more_ari-1.pdf}

\hypertarget{supplementary-figure-8-1}{%
\subsubsection{Supplementary Figure 8}\label{supplementary-figure-8-1}}

\textbf{A:} Adjusted rand index (ARI) of various clustering pipelines
using various combinations of filtering criteria. Only clustering
analyses with a number of clusters +/-1 from the real number of clusters
were considered. In general, filtering more does not result in more
accurate clustering. \textbf{B:} Percentage of cells filtered out per
subpopulation, using various combination of filters (see methods).

\hypertarget{supplementary-figure-9}{%
\section{Supplementary Figure 9}\label{supplementary-figure-9}}

\begin{verbatim}
## lm(ARI~nbClusters*dataset+norm+factor(filt.times)+factor(filt.MADs)+counts.lower+counts.higher+mt.higher+feat.lower+feat.higher+top50.lower+top50.higher+ratiodist.lower+ratiodist.higher)
\end{verbatim}

\begin{longtable}[]{@{}lrrrr@{}}
\toprule
& Estimate & Std. Error & t value & p value\tabularnewline
\midrule
\endhead
normnorm.seuratvst & 0.0069338 & 0.0003104 & 22.3415248 &
0.0000000\tabularnewline
factor(filt.times)2 & 0.0086024 & 0.0003617 & 23.7847962 &
0.0000000\tabularnewline
factor(filt.times)3 & 0.0168470 & 0.0003811 & 44.2112922 &
0.0000000\tabularnewline
factor(filt.MADs)2.5 & -0.0040702 & 0.0004199 & -9.6944075 &
0.0000000\tabularnewline
factor(filt.MADs)3 & -0.0054701 & 0.0004202 & -13.0177667 &
0.0000000\tabularnewline
factor(filt.MADs)5 & -0.0025099 & 0.0004205 & -5.9684908 &
0.0000000\tabularnewline
counts.lowerTRUE & -0.0001193 & 0.0003534 & -0.3375833 &
0.7356780\tabularnewline
counts.higherTRUE & -0.0073031 & 0.0003552 & -20.5607582 &
0.0000000\tabularnewline
mt.higherTRUE & 0.0006318 & 0.0007765 & 0.8136392 &
0.4158539\tabularnewline
feat.lowerTRUE & 0.0002074 & 0.0003494 & 0.5934348 &
0.5528917\tabularnewline
feat.higherTRUE & -0.0058113 & 0.0003506 & -16.5729653 &
0.0000000\tabularnewline
top50.lowerTRUE & -0.0000717 & 0.0003045 & -0.2354817 &
0.8138354\tabularnewline
top50.higherTRUE & -0.0000717 & 0.0003045 & -0.2354817 &
0.8138354\tabularnewline
ratiodist.lowerTRUE & -0.0009712 & 0.0003045 & -3.1895182 &
0.0014256\tabularnewline
ratiodist.higherTRUE & 0.0012337 & 0.0003045 & 4.0515679 &
0.0000509\tabularnewline
\bottomrule
\end{longtable}

\hypertarget{supplementary-figure-9-1}{%
\subsubsection{Supplementary Figure 9}\label{supplementary-figure-9-1}}

Linear regression analysis of ARI on various filtering criteria,
correcting for the effect of the number of clusters. Increasing the
number of distributions on which a cell must deviate in order to be
excluded (\texttt{filt.times}) tended to be positively associated with
ARI, while increasing the number of median absolute deviations required
to be an outlier (\texttt{filt.MADs}) was associated with a decrease in
ARI.

\hypertarget{supplementary-figure-10}{%
\section{Supplementary Figure 10}\label{supplementary-figure-10}}

\includegraphics{supp_figures_files/figure-latex/unnamed-chunk-14-1.pdf}

\includegraphics{supp_figures_files/figure-latex/unnamed-chunk-15-1.pdf}

\hypertarget{supplementary-figure-10-1}{%
\subsubsection{Supplementary Figure
10}\label{supplementary-figure-10-1}}

Impact of the set of filters on the average silhouette width of the
subpopulations (top) and on the proportion of variance in the first
components explained by clusters (bottom).

\hypertarget{supplementary-figure-11}{%
\section{Supplementary Figure 11}\label{supplementary-figure-11}}

\includegraphics{supp_figures_files/figure-latex/unnamed-chunk-16-1.pdf}

\hypertarget{supplementary-figure-11-1}{%
\subsubsection{Supplementary Figure
11}\label{supplementary-figure-11-1}}

Impact of restricting the type of features used on the ARI of the
clustering. \newpage

\hypertarget{supplementary-figure-12}{%
\section{Supplementary Figure 12}\label{supplementary-figure-12}}

\includegraphics{supp_figures_files/figure-latex/norm_r2-1.pdf}

\hypertarget{supplementary-figure-12-1}{%
\subsubsection{Supplementary Figure
12}\label{supplementary-figure-12-1}}

\textbf{A:} Correlation between library size (left) or the number of
detected features (right) and the residuals of a linear regression of
the first principal component on the subpopulation. \textbf{B:}
Proportion of the variance in the first 10 PCA components that is
explained by real subpopulations. The square-root of R-squared values is
plotted for ease of visualization.

\hypertarget{supplementary-figure-13}{%
\section{Supplementary Figure 13}\label{supplementary-figure-13}}

\includegraphics{supp_figures_files/figure-latex/norm_silwidth-1.pdf}

\hypertarget{supplementary-figure-13-1}{%
\subsubsection{Supplementary Figure
13}\label{supplementary-figure-13-1}}

Mean silhouette width of true subpopulations across the first 10
components with different normalization and scaling approaches.

\hypertarget{supplementary-figure-14}{%
\section{Supplementary Figure 14}\label{supplementary-figure-14}}

\includegraphics{supp_figures_files/figure-latex/unnamed-chunk-21-1.pdf}

\hypertarget{supplementary-figure-14-1}{%
\subsubsection{Supplementary Figure
14}\label{supplementary-figure-14-1}}

\textbf{A:} Comparison of the gene-wise proportion of variance explained
by real subpopulations based on Seurat's standard log normalization and
on \texttt{sctransform} variance-stabilizing transformation. Across 10x
datasets, there is a good agreement between the two, the correlation
ranging between 0.92 and 0.97. \textbf{B:} There is also a good
agreement between \emph{variance} and \emph{deviance} explained, with
some genes having a higher deviance explained. \textbf{C-D:}
Relationship between mean expression and the difference between the
proportion of deviance explained and the proportion of variance
explained in two datasets. Genes that have a higher proportion of the
deviance explained than of the variance explained are generally the
lowly-expressed ones.

\hypertarget{supplementary-figure-15}{%
\section{Supplementary Figure 15}\label{supplementary-figure-15}}

\begin{verbatim}
## Warning: Removed 328867 rows containing missing values (geom_path).

## Warning: Removed 328867 rows containing missing values (geom_path).
\end{verbatim}

\begin{verbatim}
## Warning: Removed 302113 rows containing missing values (geom_path).
\end{verbatim}

\begin{verbatim}
## Warning: Removed 101276 rows containing missing values (geom_path).
\end{verbatim}

\begin{verbatim}
## Warning: Removed 68502 rows containing missing values (geom_path).
\end{verbatim}

\begin{verbatim}
## Warning: Removed 94976 rows containing missing values (geom_path).
\end{verbatim}

\begin{verbatim}
## Warning: Removed 101101 rows containing missing values (geom_path).
\end{verbatim}

\begin{verbatim}
## Warning: Removed 96012 rows containing missing values (geom_path).
\end{verbatim}

\includegraphics{supp_figures_files/figure-latex/unnamed-chunk-22-1.pdf}

\hypertarget{supplementary-figure-15-1}{%
\subsubsection{Supplementary Figure
15}\label{supplementary-figure-15-1}}

Proportion of the cumulative \emph{variance} explained by real
subpulations that is retrieved through the selection. For each gene, we
compute the proportion of the variance explained by real subpopulations.
For each rank X, we sum this proportion for the X genes selected by a
given method, and divide it by the sum when selecting the X genes with
the highest variance explained. An ideal selection would therefore be a
horizontal line at 1.

\hypertarget{supplementary-figure-16}{%
\section{Supplementary Figure 16}\label{supplementary-figure-16}}

\begin{verbatim}
## Warning: Removed 328867 rows containing missing values (geom_path).

## Warning: Removed 328867 rows containing missing values (geom_path).
\end{verbatim}

\begin{verbatim}
## Warning: Removed 302113 rows containing missing values (geom_path).
\end{verbatim}

\begin{verbatim}
## Warning: Removed 101276 rows containing missing values (geom_path).
\end{verbatim}

\begin{verbatim}
## Warning: Removed 68502 rows containing missing values (geom_path).
\end{verbatim}

\begin{verbatim}
## Warning: Removed 94976 rows containing missing values (geom_path).
\end{verbatim}

\begin{verbatim}
## Warning: Removed 101101 rows containing missing values (geom_path).
\end{verbatim}

\begin{verbatim}
## Warning: Removed 96012 rows containing missing values (geom_path).
\end{verbatim}

\includegraphics{supp_figures_files/figure-latex/unnamed-chunk-23-1.pdf}

\hypertarget{supplementary-figure-16-1}{%
\subsubsection{Supplementary Figure
16}\label{supplementary-figure-16-1}}

Proportion of the cumulative \emph{deviance} explained by real
subpulations that is retrieved through the selection. For each gene, we
compute the proportion of the variance explained by real subpopulations.
As for Supplementary Figure 15, except using deviance explained.

\hypertarget{supplementary-figure-17}{%
\section{Supplementary Figure 17}\label{supplementary-figure-17}}

\includegraphics{supp_figures_files/figure-latex/unnamed-chunk-24-1.pdf}

\hypertarget{supplementary-figure-17-1}{%
\subsubsection{Supplementary Figure
17}\label{supplementary-figure-17-1}}

Proportion of the variance in the first 5 principal components that is
explained by real subpopulations according to different normalization
and selection procedures, selecting 1000 genes in each case.
